Performing Cluster Analysis on TAD boundaries & Annotations

### 2019-03-15 Scripts/evaluating_pca.Rmd
added script for results for counts and distances

# Boundaries as rows; Annotations as columns; clustering annotations

@@

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")

@@

dim(pcadat_gm12878_10kb_count) 
dim(pcadat_gm12878_10kb_count_f) 
```

@@

```{r}
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")
```

>>>>

# Clustering Annotations

@@

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")
dim(pcadat_gm12878_10kb_count)

@@

dim(pcadat_gm12878_10kb_count) 
dim(pcadat_gm12878_10kb_count_f)

colnames(pcadat_gm12878_10kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_count_f),
          keysize = 1)

#PCA
pca_gm12878_10kb_count <- prcomp(pcadat_gm12878_10kb_count_f, 
                  scale = FALSE)

##looking at the loadings
pca_gm12878_10kb_count$rotation[1:5,1:5]

##looking at the score vectors
pca_gm12878_10kb_count$x[1:5,1:5]

##screeplot
###full
std_dev <- pca_gm12878_10kb_count$sdev
pr_var <- std_dev^2
prop_varex <- pr_var/sum(pr_var)
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")
###top 10
fviz_eig(pca_gm12878_10kb_count, addlabels = TRUE)

##cumulative variance explained
plot(cumsum(prop_varex), xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")

##biplot
biplot(pca_gm12878_10kb_count, scale = 0)

##graph of the variables
fviz_pca_var(pca_gm12878_10kb_count,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("red", "white", "blue"),
             repel = TRUE     # Avoid text overlapping
             )

@@

```{r}
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")
dim(pcadat_gm12878_10kb_dist)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_dist , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_dist_f <- pcadat_gm12878_10kb_dist[, -which(colnames(pcadat_gm12878_10kb_dist) %in% nzvar)]

dim(pcadat_gm12878_10kb_dist) 
dim(pcadat_gm12878_10kb_dist_f)

colnames(pcadat_gm12878_10kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_dist_f),
          keysize = 1)

#PCA
pca_gm12878_10kb_dist <- prcomp(pcadat_gm12878_10kb_dist_f, 
                  scale = FALSE)

##looking at the loadings
pca_gm12878_10kb_dist$rotation[1:5,1:5]

##looking at the score vectors
pca_gm12878_10kb_dist$x[1:5,1:5]

##screeplot
###full
std_dev <- pca_gm12878_10kb_dist$sdev
pr_var <- std_dev^2
prop_varex <- pr_var/sum(pr_var)
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")
###top 10
fviz_eig(pca_gm12878_10kb_dist, addlabels = TRUE)

##cumulative variance explained
plot(cumsum(prop_varex), xlab = "Principal Component",
     ylab = "Cumulative Proportion of Variance Explained",
     ylim = c(0, 1), type = "b")

##biplot
biplot(pca_gm12878_10kb_dist, scale = 0)

##graph of the variables
fviz_pca_var(pca_gm12878_10kb_dist,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("red", "white", "blue"),
             repel = TRUE     # Avoid text overlapping
             )
